### 2022-08-21 chapters/chapter3.tex
update the final version

\chapter{Episodic Self-Imitation Learning with Hindsight}
\label{ch:esil}
\section{Introduction}
% \td{In this chapter, we propose episodic self-imitation learning (ESIL), which aims to address }
%------------------------------ From Main Paper ----------------------------
Reinforcement learning (RL) has been shown to be very effective in training agents within gaming environments~\cite{mnih2015human,silver2016mastering}, particularly when combined with deep neural networks~\cite{lecun2015deep,silver2016mastering,liu2017survey}. In most tasks settings that are solved by RL algorithms, reward shaping is an essential requirement for guiding the learning of the agent. However, reward shaping often requires significant quantities of domain knowledge that {are} highly task-specific~\cite{ng1999policy} and, even with careful design, can lead to undesired policies. Moreover, for complex robot manipulation tasks, the manual design of reward shaping functions to guide the learning agent becomes intractable~\cite{arulkumaran2017deep,florensa2017reverse} if even minor variations are introduced in the task. For such settings, the application of deep reinforcement learning requires algorithms that can learn from unshaped, and usually sparse, reward signals. The complicated dynamics of robot manipulation exacerbates the difficulty posed by sparse rewards, especially for on-policy RL algorithms. For example, achieving goals that require the successful execution of multiple steps over a long horizon involves high-dimensional control that must also be generalised to work across variations in the environment for each step. These aspects of robot control result in a situation where a naive RL agent so rarely receives a reward at the start of training that it cannot learn at all. A common solution in the \td{DRL} community is collecting a sufficient number of expert demonstrations, then use imitation learning to train the agent. However, in some scenarios, demonstrations are expensive to collect and the performance of a trained agent is restricted by their quantity. One solution is to use the valuable past experiences of the agent to improve training, which is particularly useful in sparse reward environments.

There are two kinds of approaches to alleviate the problems associated with sparse rewards: imitation learning and hindsight experience replay (HER)~\cite{andrychowicz2017hindsight}. First, the standard approach of imitation learning is to use supervised learning algorithms and minimise a surrogate loss with respect to an oracle. The most common form is learning from demonstrations~\cite{hester2018deep,gao2018reinforcement}. Similar techniques are applied to robot manipulation tasks~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming,james2018task}. When the demonstrations are not attainable, self-imitation learning (SIL)~\cite{oh2018self}, which uses past good experiences (episodes in which the goal is achieved), can~be used to enhance the exploration or speed up the training of the agent. {Self-imitation learning} works well in discrete control environments, such as Atari games. While being able to learn policies for continuous control tasks with dense or delayed rewards~\cite{oh2018self}, the present experiments suggest that SIL struggles when the rewards are sparse. Recently, {hindsight experience replay} has been proposed to solve such goal-oriented, sparse reward problems. The main idea of HER is that during replay, the selected transitions are sampled from state--action pairs derived from {achieved goals} that are substituted for the real goals of the task; this increases the frequency of positive rewards. {Hindsight experience replay} is used with off-policy RL algorithms, such as DQN~\cite{mnih2015human} and DDPG~\cite{lillicrap2015continuous}, for experience replay and has several extensions~\cite{schaul2015prioritized,liu2018competitive}. The present experiments show that simply applying HER with SIL does not lead to an agent capable of performing tasks from the Fetch robot environment. In summary, self-imitation learning with on-policy algorithms for tasks that require continuous control, and for which rewards are sparse, remains unsolved.

In this chapter, {episodic self-imitation learning (ESIL) for goal-oriented problems that provide only sparse rewards is proposed and combined with a state-of-the-art on-policy RL algorithm:} proximal policy optimisation (PPO). In contrast to standard SIL, which samples past good transitions from the replay buffer for imitation learning, {the proposed} ESIL adopts the entire set of current episodes {(successful or not)}, and modifies them into ``expert'' trajectories based on HER. An extra trajectory selection module is also introduced to relieve the effects of sample correlation~\cite{lee2019sample} in updating the network. Figure~\ref{fig:esil} shows the difference between naive SIL+HER and ESIL. During training by SIL+HER, a batch of transitions is sampled from the replay buffer; these are modified into ``hindsight experiences'' and used directly in self-imitation learning. In contrast, ESIL utilises the entire currently collected episodes and converts them into hindsight episodes. The trajectory selection module removes undesired transitions in the hindsight episodes. Using tasks from the OpenAI Fetch environment~\cite{plappert2018multi}, this chapter demonstrates that the proposed ESIL approach effectively trains the agent which is required to solve continuous control problems, and shows that it achieves state-of-the-art results on several tasks.

\begin{figure}[t]
\centering
\includegraphics[width=0.95\textwidth]{figures/chapter3/buffer.png}
\caption{Illustration of the difference between self-imitation learning (SIL)+hindsight experience replay (HER) and episodic self-imitation learning (ESIL).}
%% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly. PPO+ESIL utilizes entire current collected episodes and convert them into hindsight episodes. Another trajectory selection module is used to remove undesired transitions in each hindsight episodes.
\label{fig:esil}
\end{figure}

@@

\subsection{Episodic Self-Imitation Learning}
The present method aims to use episodic hindsight experiences to guide the exploration of the PPO algorithm. To this end, hindsight experiences are created from current episodes. For an episode $i$, let there be $T$ timesteps; after $T$ timesteps, a series of transitions $\tau_i =\{ (s_{t}, a_{t}, r_{t}, s_{t+1}, g, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected. If at timestep $t=T-1$, in $s_{T}$, $g^{ac}_{T} \neq g$, it implies that in this episode, the agent failed to achieve the original goal. Simply, to create hindsight experiences, {the achieved goal $g^{ac}_{T}$ in the last state $s_{T}$ is selected and considered as the modified desired goal $g^{\prime}$, i.e., $g^\prime = g^{ac}_{T}$.} Next, a new reward $r_{t}^{\prime}$ is computed under the new goal $g^\prime$. Then, {a new ``imagined'' episode is achieved}, and a new series of transitions $\tau_i^{\prime} = \{(s_{t}, a_{t}, r_{t}^{\prime},s_{t+1}, g^\prime, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected.
Then, an approach to self-imitation learning based on episodic hindsight experiences is proposed, which applies the policy updates to both hindsight and in-environment episodes. Proximal policy optimisation (PPO) is used as the base RL algorithm, which is a state-of-the-art on-policy RL algorithm. With current and corresponding hindsight experiences, a new objective function is introduced and defined as:
\begin{equation}
  \centering
  \mathcal{L} = \alpha\mathcal{L}_{PPO} + \beta\mathcal{L}_{ESIL},
  \label{eq:loss}
\end{equation}
where $\alpha$ is the weight coefficient of $\mathcal{L}_{PPO}$. In the experiments, we set $\alpha=1$ as default {to balance the contribution of $\mathcal{L}_{PPO}$ and $\mathcal{L}_{ESIL}$}. $\mathcal{L}_{PPO}$ is the loss of PPO which can be written as:
\begin{equation}
  \mathcal{L}_{PPO} = \mathcal{L}_{policy} - c\mathcal{L}_{value}
  \label{eq:ppo_loss}
\end{equation}
where $\mathcal{L}_{policy}$ is the policy loss which is parameterised by $\theta$, $\mathcal{L}_{value}$ is the value loss which is parameterised by $\psi$, and $c$ is the weight coefficient of the $\mathcal{L}_{value}$, which is set to 1 to match the default PPO setting~\cite{schulman2017proximal}. The policy loss, $\mathcal{L}_{policy}$, can be represented as:
\begin{equation}
  \centering
  \mathcal{L}_{policy} = \mathbb{E}_{s_{t}, a_{t}, g\in\Tau}\biggr[\min \left (\frac{\pi(a_{t}|s_{t}, g;\theta)}{\pi(a_{t}|s_{t}, g;\theta_{old})}A_{t}, \clip \left (\frac{\pi(a_{t}|s_{t}, g;\theta)}{\pi(a_{t}|s_{t}, g;\theta_{old})}, 1-\epsilon, 1+\epsilon \right )A_{t}\right )\biggr],
  \label{eq:policy_loss}
\end{equation}
here, $A_{t}$ is the advantage value, and can be computed as $G_{t} - V(s_{t}, g;\psi)$. $V(s_{t}, g;\psi)$ is the state value at timestep $t$ which is predicted by the critic network. $G_{t}$ is the return at timestep $t$. $\epsilon$ is the clip ratio. $\Tau$~indicates original trajectories. {The value loss is a squared error loss $\mathcal{L}_{value}=\mathbb{E}[(V(s_{t}, g;\psi) - G_{t})^{2}]$.}

@@

To prevent the agent from learning imperfect hindsight experiences, {hindsight experiences are actively selected based on returns.} With the same action, different goals may lead to different results. The proposed method only selects hindsight experiences that can achieve higher returns. The illustration of the trajectory selection module is in Figure~\ref{fig:hs_princple}. For an episodic experience and its hindsight experience, the returns of the episodic experience and its hindsight experience can be calculated, respectively. 
In a trajectory, at each timestep $t$, the return $G_{t}$ can be calculated by $G_{t} = G(s_t, g) = r_{t} + \gamma r_{t+1} + \gamma^{2} r_{t+2} + ... + \gamma^{T-t-1} r_{T-1}$. Then, for a trajectory $\tau_{i}$, we have $\left\{G_{0}^{i}, G_{1}^{i}, G_{2}^{i}, ..., G_{T-1}^{i}\right\}$.

For the hindsight experiences, similarly, the return $G_{t}'$ for each timestep $t$,~with respect to the hindsight goal $g^\prime$, can be calculated. Based on the modified trajectory $\tau_{i}^{\prime}$ with the same length of $\tau_{i}$, we therefore have the returns $\left\{G_{0}^{i^\prime}, G_{1}^{i^\prime}, G_{2}^{i^\prime}, ..., G_{T-1}^{i^\prime}\right\}$. During training, the~hindsight experiences with higher returns are used for self-imitation learning. The rest of the hindsight experiences will be \td{assumed to be of lower importance} and ignored.
Then, Equation~\eqref{eq:imitaion} can be rewritten as:
\begin{equation}
  \centering
  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'},g\in\Tau}\biggr[\log\pi_{\theta}\left(a_{t}|s_{t}, g^\prime\right)\mathcal{F}\left(s_{t}, g, g^\prime\right)\biggr],
  \label{eq:final_loss}
\end{equation}
where $\mathcal{F}\left(s_{t}, g, g^{\prime}\right)$ is the trajectory selection module. The selection function can be expressed as
\begin{equation}
  \centering
  \mathcal{F}\left(s_{t}, g, g^\prime \right)=\mathds{1}\left[G(s_{t}, g^\prime) > G(s_{t}, g)\right],
\end{equation}
here, {$\mathds{1}(\cdot)$ is the unit step function}. Consider the OpenAI Fetch environment as an example. For~a failed trajectory, the rewards $r_{t}$ are $\{-1,-1,\cdots, -1\}$. {The desired goal is modified} to construct a new hindsight trajectory, and the new rewards $r_{t}^{\prime}$ become $\{-1,-1,\cdots, 0\}$. Then,  {$G$ and $G^{\prime}$ can be calculated~separately.}
\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/chapter3/hgs.png}
  \caption{A simplified illustration of trajectory selection. Blue trajectories indicate original experiences. Orange trajectories indicate hindsight experiences. Solid trajectories in the hindsight experiences are selected by the trajectory selection module of ESIL with new ``imagined'' goals.}
  \label{fig:hs_princple}
\end{figure}

From a goal perspective, episodic self-imitation learning (ESIL) tries to explore (desired) goals to get positive returns.~It can be viewed as a form of multi-task learning, {because ESIL has two objective functions to be optimised jointly.}~It is also related to self-imitation learning (SIL)~\cite{oh2018self}.~However,~ the~difference is that SIL uses $\left(G-V(s;\theta)\right)_{+}$ on past experiences to learn to choose the action chosen in the past in a given state, rather than goals. The full description of ESIL can be found in Algorithm~\ref{alg}.
% the algorithm
\begin{algorithm}[t!]
  \caption{Proximal Policy Optimisation (PPO) with Episodic Self-Imitation Learning (ESIL)}
  \label{alg}
  \begin{algorithmic}[1]
    \REQUIRE an actor network $\pi(s, g;\theta)$, a critic network $V(s, g;\psi)$, the maximum steps $T$ of an episode, a reward function $\mathcal{R}$, number of iterations $N$, number of episodes for each iteration $M$, weight coefficient of critic loss term $c$, weight coefficient of PPO term $\alpha$, adaptive coefficient $\beta$  
    %\STATE Initialize the parameters $\theta$ and $\psi$ of both actor and critic networks
    \FOR{$\mathrm{iteration} = 1, 2, \cdots N$}
    \STATE $\Tau \leftarrow \varnothing$, $\Tau^{\prime} \leftarrow \varnothing$ 
    \FOR{$\mathrm{episode} = 1, 2, ..., M$}
    \STATE $\tau \leftarrow \varnothing$
    \FOR{$t = 0, 1, \cdots, T-1$}
    \STATE Sample an action $a_{t}$ using the actor network $\pi(s_{t}, g;\theta)$
    \STATE Execute the action $a_{t}$ and observe a new state $s_{t+1}$
    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}, s_{t+1}, g^{ac}_{t+1}\right)$ in $\tau$
    \ENDFOR

@@

\section{Experiments}
\label{sec:experiment}
The proposed method is evaluated on several multi-goal environments, including the Empty Room environment and the OpenAI Fetch environment (see Figure~\ref{fig:env}). The Empty Room environment is a toy example, and has discrete action spaces. In the Fetch environment, there are four robot tasks with continuous action spaces. To obtain a comprehensive comparison between the proposed method and other baselines, suitable baselines are selected for different environments. In addition, ablation studies of the trajectory selection module are also performed.
\begin{figure}[t]
\minipage{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room.png}
  ({a}) Empty Room         
\endminipage\hfill
\minipage{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach.png}
  ({b}) Reach 
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push.png}
  ({c}) Push      
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick.png}
  ({d}) Pick$\&$Place
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide.png}
  ({e}) Slide     
\endminipage
\caption{Evaluation environments. ({a}) is the Empty Room environment, in which a yellow circle indicates the position of the agent and a red star represents a target position. ({b}--{e}) are the tasks in the Fetch robot environment. The red spot represents a target position.}
\label{fig:env}
\end{figure}

\subsection{Environments}
\textbf{{Empty Room (grid-world) environment}}:
The Empty Room environment is a simple grid-world environment. The agent is placed in an $11\times11$ grid, representing the room. The goal of the agent is to reach a target position in the room. The start position of the agent is at the left upper corner of the room, and the target position is randomly selected  within the room. When the agent chooses an action that would fall outside the grid area, the agent will stay at the current position. The length of each episode is 32. 
The desired goal, $g$, is a two-dimensional grid coordinate, which represents the target position. The achieved goal, $g^{ac}_{t}$, is also a two-dimensional coordinate, which represents the current position of the agent, and finally, the observation is a two-dimensional coordinate which represents the current position of the agent. 
The agent has five actions: left, right, up, down, and stay; the agent executes a random action with a probability of 0.2. The agent can get $+1$ as a reward only when $g^{ac}_{t+1}=g$, otherwise, it gets a reward of $0$.

>>>>

\chapter{Episodic Self-Imitation Learning with Hindsight}
\label{ch:esil}
\section{Introduction}
% \td{In this chapter, we propose episodic self-imitation learning (ESIL), which aims to address }
%------------------------------ From Main Paper ----------------------------
Reinforcement learning (RL) has been shown to be very effective in training agents within gaming environments~\cite{mnih2015human,silver2016mastering}, particularly when combined with deep neural networks~\cite{lecun2015deep,silver2016mastering,liu2017survey}. In most tasks settings that are solved by RL algorithms, reward shaping is an essential requirement for guiding the learning of the agent. However, reward shaping often requires significant quantities of domain knowledge that {are} highly task-specific~\cite{ng1999policy} and, even with careful design, can lead to undesired policies. Moreover, for complex robot manipulation tasks, the manual design of reward shaping functions to guide the learning agent becomes intractable~\cite{arulkumaran2017deep,florensa2017reverse} if even minor variations are introduced in the task. For such settings, the application of deep reinforcement learning requires algorithms that can learn from unshaped, and usually sparse, reward signals.

The complicated dynamics of robot manipulation exacerbates the difficulty posed by sparse rewards, especially for on-policy RL algorithms. For example, achieving goals that require the successful execution of multiple steps over a long horizon involves high-dimensional control that must also be generalised to work across variations in the environment for each step. These aspects of robot control result in a situation where a naive RL agent so rarely receives a reward at the start of training that it cannot learn at all. A common solution in the DRL community is collecting a sufficient number of expert demonstrations, then use imitation learning to train the agent. However, in some scenarios, demonstrations are expensive to collect and the performance of a trained agent is restricted by their quantity. One solution is to use the valuable past experiences of the agent to improve training, which is particularly useful in sparse reward environments.

There are two types of approach to alleviate the problems associated with sparse rewards: imitation learning and hindsight experience replay (HER)~\cite{andrychowicz2017hindsight}. First, the standard approach of imitation learning is to use supervised learning algorithms and minimise a surrogate loss with respect to an oracle. The most common form is learning from demonstrations~\cite{hester2018deep,gao2018reinforcement}. Similar techniques are applied to robot manipulation tasks~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming,james2018task}. When the demonstrations are not attainable, self-imitation learning (SIL)~\cite{oh2018self}, which uses past good experiences (episodes in which the goal is achieved), can~be used to enhance the exploration or speed up the training of the agent. {Self-imitation learning} works well in discrete control environments, such as Atari games. While being able to learn policies for continuous control tasks with dense or delayed rewards~\cite{oh2018self}, the present experiments suggest that SIL struggles when the rewards are sparse.

Recently, {hindsight experience replay} has been proposed to solve such goal-oriented, sparse reward problems. The main idea of HER is that during replay, the selected transitions are sampled from state--action pairs derived from {achieved goals}. This increases the frequency of positive rewards. {Hindsight experience replay} is used with off-policy RL algorithms, such as DQN~\cite{mnih2015human} and DDPG~\cite{lillicrap2015continuous}, for experience replay and has several extensions~\cite{schaul2015prioritized,liu2018competitive}. The present experiments show that simply applying HER with SIL does not lead to an agent capable of performing tasks from the Fetch robot environment. In summary, self-imitation learning with on-policy algorithms for tasks that require continuous control, and for which rewards are sparse, remains unsolved.

\begin{figure}[t]
\centering
\includegraphics[width=0.95\textwidth]{figures/chapter3/buffer.png}
\caption[Illustration between SIL+HER and ESIL.]{Illustration of the difference between self-imitation learning (SIL)+hindsight experience replay (HER) and episodic self-imitation learning (ESIL).}
%% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly. PPO+ESIL utilizes entire current collected episodes and convert them into hindsight episodes. Another trajectory selection module is used to remove undesired transitions in each hindsight episodes.
\label{fig:esil}
\end{figure}

In this chapter, {episodic self-imitation learning (ESIL) for goal-oriented problems that provide only sparse rewards is proposed and combined with a state-of-the-art on-policy RL algorithm:} proximal policy optimisation (PPO). In contrast to standard SIL, which samples past good transitions from the replay buffer for imitation learning, {the proposed} ESIL adopts the entire set of current episodes {(successful or not)}, and modifies them into ``expert'' trajectories based on HER. An extra trajectory selection module is also introduced to relieve the effects of sample correlation~\cite{lee2019sample} in updating the network. Figure~\ref{fig:esil} shows the difference between naive SIL+HER and ESIL. We further provide an illustration (Figure~\ref{fig:type_buffer}) to distinguish between the different types of buffer used in each algorithm in this chapter. Figure~\ref{fig:type_buffer}a shows the structure of an \textit{episodic buffer}, which is used in the proposed ESIL. In the episodic buffer, each entry stores a complete episode (i.e., trajectory) with the same length $T$. In each update iteration, all episodes in this buffer will be used for training. After each update iteration, the stored episodes in this buffer will be removed. The episodic buffer only stores the episodes that are collected by the current policy of the agent. Figure~\ref{fig:type_buffer}b shows the structure of a standard \textit{replay buffer}, which is used by SIL and SIL+HER in this chapter. Instead of storing a complete episode in each entry, the replay buffer stores a transition in each entry. In each update iteration, batch-sized transitions are sampled for training. After each update iteration, the buffer will not be cleaned. The replay buffer will remove the oldest transition when it is full and a new transition needs to be stored. Figure~\ref{fig:type_buffer}c shows the structure of an \textit{episodic replay buffer}, which is used by DDPG+HER in this chapter. Similar to episodic buffer, the episodic replay buffer stores a complete episode with the same length in each entry. In each update iteration, batch-sized episodes will be uniformly sampled. Then, one transition will be uniformly sampled from each selected episode to form batch-sized transitions for training. After each update iteration, the buffer will not be cleaned. The episodic replay buffer will remove the oldest episode when it is full and a new episode needs to be stored.

\begin{figure}[h]
\centering
\includegraphics[width=0.95\textwidth]{figures/chapter3/buffer_type.pdf}
\caption[Illustration of different types of buffer.]{Illustration of different types of buffers: (a) is the episodic buffer, which is used in ESIL; (b) is the standard replay buffer, which is used in SIL and SIL+HER; (c) is the episodic replay buffer, which is used in HER. Here, $N$ is the number of samples or episodes, T is the length of each episode, D is dimension of the stored elements (e.g., state $s$).}
%% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly. PPO+ESIL utilizes entire current collected episodes and convert them into hindsight episodes. Another trajectory selection module is used to remove undesired transitions in each hindsight episodes.
\label{fig:type_buffer}
\end{figure}

During training by SIL+HER, a batch of transitions is sampled from the replay buffer; these are modified into ``hindsight experiences'' and used directly in self-imitation learning. In contrast, ESIL utilises the entire currently collected episodes and converts them into hindsight episodes. The trajectory selection module removes undesired transitions in the hindsight episodes. Using tasks from the OpenAI Fetch environment~\cite{plappert2018multi}, this chapter demonstrates that the proposed ESIL approach effectively trains the agent which is required to solve continuous control problems, and shows that it achieves state-of-the-art results on several tasks.

@@

\subsection{Episodic Self-Imitation Learning}
The method aims to use episodic hindsight experiences to guide the exploration of the PPO algorithm. To this end, hindsight experiences are created from current episodes. For an episode $i$, let there be $T$ timesteps; after $T$ timesteps, a series of transitions $\tau_i =\{ (s_{t}, a_{t}, r_{t}, s_{t+1}, g, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected. If at timestep $t=T-1$, in $s_{T}$, $g^{ac}_{T} \neq g$, it implies that in this episode, the agent failed to achieve the original goal. Simply, to create hindsight experiences, {the achieved goal $g^{ac}_{T}$ in the last state $s_{T}$ is selected and considered as the modified desired goal $g^{\prime}$, i.e., $g^\prime = g^{ac}_{T}$.} Next, a new reward $r_{t}^{\prime}$ is computed under the new goal $g^\prime$. Then, {a new ``imagined'' episode is created}, and a new series of transitions $\tau_i^{\prime} = \{(s_{t}, a_{t}, r_{t}^{\prime},s_{t+1}, g^\prime, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected.
Then, an approach to self-imitation learning based on episodic hindsight experiences is proposed, which applies the policy updates to both hindsight and in-environment episodes. Proximal policy optimisation (PPO) is used as the base RL algorithm, which is a state-of-the-art on-policy RL algorithm. With current and corresponding hindsight experiences, a new objective function is introduced and defined as:
\begin{equation}
  \centering
  \mathcal{L} = \alpha\mathcal{L}_{PPO} + \beta\mathcal{L}_{ESIL},
  \label{eq:loss}
\end{equation}
where $\alpha$ is the weight coefficient of $\mathcal{L}_{PPO}$. In the experiments, we set $\alpha=1$ as default {to balance the contribution of $\mathcal{L}_{PPO}$ and $\mathcal{L}_{ESIL}$}. $\mathcal{L}_{PPO}$ is the loss of PPO which can be written as:
\begin{equation}
  \mathcal{L}_{PPO} = \mathcal{L}_{policy} - c\mathcal{L}_{value}
  \label{eq:ppo_loss}
\end{equation}
where $\mathcal{L}_{policy}$ is the policy loss which is parameterised by $\theta$, $\mathcal{L}_{value}$ is the value loss which is parameterised by $\psi$, and $c$ is the weight coefficient of the $\mathcal{L}_{value}$, which is set to 1 to match the default PPO setting~\cite{schulman2017proximal}. The policy loss, $\mathcal{L}_{policy}$, can be represented as:
\begin{equation}
  \centering
  \mathcal{L}_{policy} = \mathbb{E}_{s_{t}, a_{t}, g\in\Tau}\biggr[\min \left (\frac{\pi(a_{t}|s_{t}, g;\theta)}{\pi(a_{t}|s_{t}, g;\theta_{old})}A_{t}, \clip \left (\frac{\pi(a_{t}|s_{t}, g;\theta)}{\pi(a_{t}|s_{t}, g;\theta_{old})}, 1-\epsilon, 1+\epsilon \right )A_{t}\right )\biggr],
  \label{eq:policy_loss}
\end{equation}
here, $A_{t}$ is the advantage value, and can be computed as $G_{t} - V(s_{t}, g;\psi)$. $V(s_{t}, g;\psi)$ is the state value at timestep $t$ which is predicted by the critic network. $G_{t}$ is the return at timestep $t$. $\epsilon$ is the clip ratio. $\Tau$~indicates original trajectories. {The value loss is a squared error loss $\mathcal{L}_{value}=\mathbb{E}[(V(s_{t}, g;\psi) - G_{t})^{2}]$.}

@@

To prevent the agent from learning imperfect hindsight experiences, {hindsight experiences are actively selected based on returns.} With the same action, different goals may lead to different results. The proposed method only selects hindsight experiences that can achieve higher returns. 
%The illustration of the trajectory selection module is in Figure~\ref{fig:hs_princple}. 
%\td{An original trajectory can be divided into two categories: 1) the agent never reaches the goal and 2) the agent has reached its original before. Figure~\ref{fig:hs_princple} A shows the case that the agent never reaches the goal. In this case, the }

For an episodic experience and its hindsight experience, the returns of the episodic experience and its hindsight experience can be calculated, respectively. 
In a trajectory, at each timestep $t$, the return $G_{t}$ can be calculated by $G_{t} = G(s_t, g) = r_{t} + \gamma r_{t+1} + \gamma^{2} r_{t+2} + ... + \gamma^{T-t-1} r_{T-1}$. Then, for a trajectory $\tau_{i}$, we have $\left\{G_{0}^{i}, G_{1}^{i}, G_{2}^{i}, ..., G_{T-1}^{i}\right\}$. For the hindsight experiences, similarly, the return $G_{t}'$ for each timestep $t$,~with respect to the hindsight goal $g^\prime$, can be calculated. Based on the modified trajectory $\tau_{i}^{\prime}$ with the same length of $\tau_{i}$, we therefore have the returns $\left\{G_{0}^{i^\prime}, G_{1}^{i^\prime}, G_{2}^{i^\prime}, ..., G_{T-1}^{i^\prime}\right\}$.

The illustration of the trajectory selection module is shown in Figure~\ref{fig:hs_princple}. An original trajectory collected by an agent can be divided into two categories: 1) the agent never reaches the goal and 2) the agent has reached its original goal before. Figure~\ref{fig:hs_princple}A shows the case where the agent never reaches the goal. In this case, the original experiences never contain a non-negative reward. However, the hindsight experience will have at least one transition that has a non-negative reward (e.g., the final step). According to the rules for calculating the returns, all hindsight experiences (solid orange trajectory) have higher returns than the original experiences (solid blue trajectory). Therefore, in this case, all hindsight experiences can be selected to benefit the training. Figure~\ref{fig:hs_princple}B demonstrates the case that the agent has achieved its original goal in the third timestep and thus received a non-negative reward. However, for the hindsight experience, it only achieves a non-negative reward in the end of the trajectory. Therefore, in the first three steps, the original experiences have higher returns than the hindsight experiences (dashed orange trajectory), which means that the original experiences have better quality than the hindsight experiences, and these hindsight experiences are ignored during training. For the remainder of the episode of Figure~\ref{fig:hs_princple}B, the hindsight experiences (solid orange trajectory) have higher returns than the original ones, and thus these hindsight experiences are selected to accelerate training.

During training, the~hindsight experiences with higher returns are used for self-imitation learning. The rest of the hindsight experiences will be assumed to be of lower importance and ignored.
Then, Equation~\eqref{eq:imitaion} can be rewritten as:
\begin{equation}
  \centering
  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'},g\in\Tau}\biggr[\log\pi_{\theta}\left(a_{t}|s_{t}, g^\prime\right)\mathcal{F}\left(s_{t}, g, g^\prime\right)\biggr],
  \label{eq:final_loss}
\end{equation}
where $\mathcal{F}\left(s_{t}, g, g^{\prime}\right)$ is the trajectory selection module. The selection function can be expressed as
\begin{equation}
  \centering
  \mathcal{F}\left(s_{t}, g, g^\prime \right)=\mathds{1}\left[G(s_{t}, g^\prime) > G(s_{t}, g)\right],
\end{equation}
here, {$\mathds{1}(\cdot)$ is the unit step function}. Consider the OpenAI Fetch environment as an example. For~a failed trajectory, the rewards $r_{t}$ are $\{-1,-1,\cdots, -1\}$. {The desired goal is modified} to construct a new hindsight trajectory, and the new rewards $r_{t}^{\prime}$ become $\{-1,-1,\cdots, 0\}$. Then,  {$G$ and $G^{\prime}$ can be calculated~separately.}
\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/chapter3/hgs.png}
  \caption[A simplified illustration of trajectory selection.]{A simplified illustration of trajectory selection. Blue trajectories indicate original experiences. Orange trajectories indicate hindsight experiences. Solid trajectories in the hindsight experiences are selected by the trajectory selection module of ESIL with new ``imagined'' goals.}
  \label{fig:hs_princple}
\end{figure}

From a goal perspective, episodic self-imitation learning (ESIL) tries to explore (desired) goals to get positive returns.~It can be viewed as a form of multi-task learning, {because ESIL has two objective functions to be optimised jointly.}~It is also related to self-imitation learning (SIL)~\cite{oh2018self}. However, the difference is that SIL uses $\left(G-V(s;\theta)\right)_{+}$ on past experiences to learn to choose the action chosen in the past in a given state, rather than goals. The full description of ESIL can be found in Algorithm~\ref{alg}.
% the algorithm
\begin{algorithm}[t!]
  \caption{Proximal Policy Optimisation (PPO) with Episodic Self-Imitation Learning (ESIL)}
  \label{alg}
  \begin{algorithmic}[1]
    \REQUIRE an actor network $\pi(s, g;\theta)$, a critic network $V(s, g;\psi)$, the maximum steps $T$ of an episode, a reward function $\mathcal{R}$, number of iterations $N$, number of episodes for each iteration $M$, weight coefficient of critic loss term $c$, weight coefficient of PPO term $\alpha$, adaptive coefficient $\beta$  
    %\STATE Initialize the parameters $\theta$ and $\psi$ of both actor and critic networks
    \FOR{$\mathrm{iteration} = 1, 2, \cdots N$}
    \STATE $\Tau \leftarrow \varnothing$, $\Tau^{\prime} \leftarrow \varnothing$ 
    \FOR{$\mathrm{episode} = 1, 2, ..., M$}
    \STATE $\tau \leftarrow \varnothing$
    \FOR{$t = 0, 1, \cdots, T-1$}
    \STATE Sample an action $a_{t}$ using the actor network $\pi(s_{t}, g;\theta)$
    \STATE Execute the action $a_{t}$ and observe a new state $s_{t+1}$
    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}, s_{t+1}, g, g^{ac}_{t+1}\right)$ in $\tau$
    \ENDFOR

@@

\section{Experiments}
\label{sec:experiment}
The proposed method is evaluated on several multi-goal environments, including the Empty Room environment and the OpenAI Fetch environment (see Figure~\ref{fig:env}). The Empty Room environment is a toy example, and has discrete action spaces. In the Fetch environment, there are four robot tasks with continuous action spaces. To obtain a comprehensive comparison between the proposed method and other baselines, suitable baselines are selected for different environments. In addition, ablation studies of the trajectory selection module are also performed.
\begin{figure}[t]
\minipage{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room.png}
  ({a}) Empty Room         
\endminipage\hfill
\minipage{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach.png}
  ({b}) Reach 
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push.png}
  ({c}) Push      
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick.png}
  ({d}) Pick$\&$Place
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide.png}
  ({e}) Slide     
\endminipage
\caption[Evaluation environments of ESIL.]{Evaluation environments. ({a}) is the Empty Room environment, in which a yellow circle indicates the position of the agent and a red star represents a target position. ({b}--{e}) are the tasks in the Fetch robot environment. The red spot represents a target position.}
\label{fig:env}
\end{figure}

\subsection{Environments}
\textbf{{Empty Room (grid-world) environment}}:
The Empty Room environment is a simple grid-world environment. The agent is placed in an $11\times11$ grid, representing the room. The goal of the agent is to reach a target position in the room. The start position of the agent is at the left upper corner of the room, and the target position is randomly selected within the room at the beginning of each episode~\cite{rauber2018hindsight}. When the agent chooses an action that would fall outside the grid area, the agent will stay at the current position. The length of each episode is 32. 
The desired goal, $g$, is a two-dimensional grid coordinate, which represents the target position. The achieved goal, $g^{ac}_{t}$, is also a two-dimensional coordinate, which represents the current position of the agent, and finally, the observation is a two-dimensional coordinate which represents the current position of the agent. 
The agent has five actions: left, right, up, down, and stay; the agent executes a random action with a probability of 0.2. The agent can get $+1$ as a reward only when $g^{ac}_{t+1}=g$, otherwise, it gets a reward of $0$.
